---
title: "Homework 1"
author: "David Blumenstiel"
date: "8/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
```

## Problem set 1

### (1) Calculate the dot product u.v where u = [0.5; 0.5] and v = [3;-4]

```{r}
u <- c(0.5, 0.5)
v <- c(3, -4)

i = 0
buff = c()
while (i < length(u)) {
   
  i = i+1
  mult = u[i]*v[i]
  buff[[i]] <- mult
  
}

print(sum(buff))

#or if you wana use a package
#print(dot(u,v))  
  
```

The product of u.v is -0.5

### (2) What are the lengths of u and v? Please note that the mathematical notion of the length of a vector is not the same as a computer science definition.
```{r}
ulen <-(0.5**2 + 0.5**2)**.5
vlen <- (3**2 + (-4)**2)**.5

print(ulen)
print(vlen)
```

The length of the 'u' is about 0.707, and the length of 'v' is 5.

### (3) What is the linear combination: 3u - 2v?

```{r}
u3 <- 3 * u
v2 <- 2 * v
print(u3)
print(v2)

print(u3 + v2)
print(sum(u3 + v2))
```
The linear combination is 1

### (4) What is the angle between u and v


```{r}
costheta <- dot(u,v)/(ulen * vlen)

theta <- acos(costheta)
print(180 * theta / pi)
```
The angle between the two vectors is 98.13 deg.

## Problem set 2

Set up a system of equations with 3 variables and 3 constraints and solve for x. Please
write a function in R that will take two variables (matrix A & constraint vector b) and
solve using elimination. Your function should produce the right answer for the system of
equations for any 3-variable, 3-equation system. You don't have to worry about degenerate
cases and can safely assume that the function will only be tested with a system of equations
that has a solution. Please note that you do have to worry about zero pivots, though. Please
note that you should not use the built-in function solve to solve this system or use matrix
inverses. The approach that you should employ is to construct an Upper Triangular Matrix
and then back-substitute to get the solution. Alternatively, you can augment the matrix
A with vector b and jointly apply the Gauss Jordan elimination procedure.

let's start by making a function and testing it out
```{r}
set.seed("1234567890")

A <- matrix(rexp(9), ncol=3)    # Makes a random 3X3 matrix
b <- matrix(rexp(3), ncol = 1)  # A constraint vector

#This is a function which uses the elimination method to 3X3 solve systems of equations
eliminate3 <- function(A, b) {
  
  
  #Gets seperate rows
  row1 = A[1,]
  row2 = A[2,]
  row3 = A[3,]
  
  
  #This all makes an upper triangular matrix by pivoting to 0 all else
  if (row2[1] != 0) {
  
    b[2] = b[2] - (row2[1] / row1[1] * b[1])
    buff = row2[1] / row1[1] * row1
  
    row2 = row2 - buff
  
  }
  
  
  if (row3[1] != 0) {
  
    b[3] = b[3] - (row3[1] / row1[1] * b[1])
    buff = row3[1] / row1[1] * row1
  
    row3 = row3 - buff
  
  }
  
  if (row3[2] != 0 & row2[2] !=0) {
  
    b[3] = b[3] - (row3[2] / row2[2] * b[1])
    buff = row3[2] / row2[2] * row2
  
    row3 = row3 - buff
  
  } else if (row3[2] != 0 & row1[2] != 0) {
    
    b[3] = b[3] - (row3[2] / row1[2] * b[1])
    buff = row3[2] / row1[2] * row2
    
    row3 = row3 - buff
    
  }
  
  x <- matrix(ncol = 3, nrow = 1) # A blank vector for the solutions
  
  #Uses the upper triangular matrix to solve for the variables
  
  x[3] = b[3]/row3[3]
  x[2] = (b[2] - (row2[3] * x[3])) / row2[2]
  x[1] = (b[1] - (row1[3] * x[3]) - (row1[2] * x[2])) / row1[1]
  
    print(row1)
    print(row2)
    print(row3)
    return(x)
    
}

#Now let's test it out with a random matrix and constraint vector

print(A)
print(b)

print(eliminate3(A, b))
```

Below, the elimination function is used to solve the system given in the homework
```{r}
A <- matrix(c(1,2,-1,1,-1,-2,3,5,4), ncol = 3)
print(A)

b <- matrix(c(1,2,6), ncol = 1)
print(b)

eliminate3(A,b)
```

My function yields a solution vector of -1.42, - 0.30, and 0.91.  It's very close to the actual solution, and after looking through the code for a while, I'm still not sure why it isn't the exact solution.  